---
title: " Linear Mixed-Effects Model Summary"
output:
  html_document: default
---

Let's investigate our data to see what the numbers can tell us about eating fish and eating sugar and how both relate to happiness scores and incidences of anxiety & depressive disorders across many countries around the world.

*Remember: not all countries were represented in each data set collected, and when refining some were lost. This means not all countries are accounted for in this investigation, though 150 countries is no small number*

------------------------------------------------------------------------

## Initial imports

**First we'll import the libraries we're going to use.**

```{r}
shh <- suppressPackageStartupMessages
shh(library(tidyverse))
shh(library(broom.mixed))
shh(library(lme4))
shh(library(ggeffects))
```

#### Next we'll import our data.

*You'll notice that the data has already been cleaned and arranged in a way that makes the following operations straightforward. The original data files are available from their sources (linked on the home page), or in the github repository.*

```{r}
data <- tibble(read_csv('../../data/happy_sad_sugar_fish.csv')) |> 
  arrange(country)
data
```

------------------------------------------------------------------------

## Modeling the data

Linear mixed effects regression models are used when not all observations are independent incorporating variables that have both fixed effects and random effects.

In our case, the **fixed effects predictors** are `fish_consumption`, `sugar_consumption`, and `year`. The **random effects predictor** is `country`. The target variable is either the `happiness_score` or the `sadness_score`.

We'll use the `lmer` package to evaluate.

```{r}
mixed_model <- lmer(
  happiness_score ~ year + sugar_g_per_person_per_day + 
    fish_kg_per_person_per_year + (1 | country), 
  data = data)
summary(mixed_model)
```

#### Random Effects

```         
Random effects:
 Groups   Name        Variance Std.Dev.
 country  (Intercept) 91.97    9.590   
 Residual             14.87    3.856   
Number of obs: 1600, groups:  country, 150
```

The variance between countries ($91.97$) is much larger than the residual variance, suggesting that **there are substantial differences in baseline happiness between countries**.

-   this means that the difference in country accounts for much more of the variation in happiness score than the unexplained variation within countries.

-   the ratio $91.97/(91.97 + 14.87) = 0.86$ suggests that around $86%$ of the variation in happiness scores (after accounting for the fixed effects) is explained by differences in baseline happiness by country.

#### Fixed Effects

```         
Fixed effects:
                             Estimate Std. Error t value
(Intercept)                 -0.048114  56.500736  -0.001
year                         0.025223   0.028171   0.895
sugar_g_per_person_per_day   0.017764   0.004474   3.971
fish_kg_per_person_per_year  0.113243   0.036758   3.081
```
We gain some interesting insight when looking at the fixed effects. There is a positive association between `year`, `sugar_consumption`, and `fish_consumption`. 

The $t-values$ (calculated as $t = Estimate / Std. Error$) will tell us what is statistically significant and what is not. As a rule of thumb, when using mixed effects models:

  * \$t$\ > 2 suggests statistical significance at approximately the $p = 0.05$ level.
  
  * \$t$\ > 2.6 suggests significance at approximately the 0.01 level.
  
  * \$t$\ > 3.3 suggests significance at approximately the 0.001 level.

Based on the $t$ values, we conclude that `year` is **not** statistically significant, `fish_consumption` is **very** statistically significant, and `sugar_consumption` is **highly** statistically significant. 

```{r}
# Extract predictions while varying one predictor at a time
effects_year  <- ggpredict(mixed_model, terms = "year")  
effects_sugar <- ggpredict(mixed_model, terms = "sugar_g_per_person_per_day")
effects_fish  <- ggpredict(mixed_model, terms = "fish_kg_per_person_per_year")

# Plot each effect separately
plot1 <- ggplot(effects_year, aes(x = x, y = predicted)) +
  geom_line(color = "blue") +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +
  labs(title = "Effect of Year on\nHappiness Score", x = "Year", y = "Predicted Happiness Score")

plot2 <- ggplot(effects_sugar, aes(x = x, y = predicted)) +
  geom_line(color = "red") +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, fill = "red") +
  labs(title = "Effect of Sugar Consumption\non Happiness Score", x = "Sugar (g per person/day)", y = "Predicted Happiness Score")

plot3 <- ggplot(effects_fish, aes(x = x, y = predicted)) +
  geom_line(color = "orange") +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, fill = "green") +
  labs(title = "Effect of Fish Consumption\non Happiness Score", x = "Fish (kg per person/year)", y = "Predicted Happiness Score")

# Arrange plots in a grid
library(patchwork)
plot1 + plot2 + plot3

```
